Covid-19 Fields Survey Data Protocol

### 2023-01-13 protocol/Background_and_methods_Jan_12.qmd
updated protocol

---
title: "Covid Fields Data Protocol"
author:
  - name: Ariel Mundo Ortiz
    orcid: 0000-0002-6014-4538
    email: ariel.mundo.ortiz@umontreal.ca
    affiliations:
      - name: École de santé publique - Université de Montréal
        address: 7101 Park Avenue
        city: Montreal
        state: QC
        postal-code: H3N 1X9
format:
  pdf:
    link-citations: true
    number-sections: true
editor: visual
bibliography: bibliography/refs.bib
keep-tex: true
csl: bibliography/jama.csl
knitr:
  opts_chunk:
    dev: "ragg_png"
---

@@

Studies that analyze geographical variations in vaccine uptake can help inform public health decision-makers to design policies to that are aimed at addressing vaccination disparities. In this regad, previous geographical (spatial) analyses of vaccination rates have shown that variations in vaccine uptake can occur within small governamental administrative units (e.g., counties in the case of the US) [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that geographical analyses can be predictive of booster uptake patterns [@wood2022].

In Canada, studies that have used a spatial approach to analyze vaccine uptake have shown disparities in vaccination rates across low and high income neighborhoods in the city of Toronto [@choi2021], among adolescents from deprived neighborhoods in the city of Montreal [@mckinnon2021], and highlighted disparities in vaccination status depending on age, income, and ethnic origin in all of the Canadian provinces[@guay2022]. However, to the best of our knowledge, there are no studies that have analyzed vaccination status within a province at the district/rural municipality level, which can be useful to identify inequalities over these geographical areas, thus providing a dissagregated view that can help understand the barriers for vaccine delivery in the case of visible minorities, which have been disproportionately impacted in the pandemic [@hussain2022].

@@

This study will examine self-reported COVID-19 vaccination status in the province of Ontario in order to determine the influence in vaccination status due to socio-economic (e.g., ethnic origin, age, income) and geographical factors (at the municipal level).

@@

| Variable             | Values                                                                                                                          |
|:---------------------|---------------------------------------------------------------------------------------------------------------------------------|
| Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
| Income bracket (CAD) | \<15,000, 15,000-24,999, 25,000-39,999, 40,000-59,999, 60,000-89,999, \>90,000                                                  |
| Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, other |
| Employment status    | yes, no                                                                                                                         |

@@

We selected two of the questions that were asked in the survey regarding vaccination status:

@@

-   (If answered "yes" to the previous question) "Have you received the second dose of the COVID vaccine?" with possible answers "yes" and "no"

@@

The original dataset contained 39,029 entries (where each entry corresponded to a set of answers provided by a unique respondent). Following a preliminary analysis to identify the missing rates across the different answers within each entry, it was identified that many of the answers had high missing rates (\>80%) (note that the graph will be in the Appendix). Therefore, the dataset was cleaned in order to contain only the independent variables of interest with the lowest missing rates (@tbl-covariates) and the dependent variables.

@@

@@

Differences between the clean dataset and the 2016 Census data for Ontario regarding proportions of age groups, income, and ethnicity/race were identified. Therefore, the proportions of each of these variables were corrected using an iterative proportional fitting procedure (also known as *raking*) [@deming1940] in R using the `survey` package. Proportions for the correction were obtained from the 2016 Census Data for Ontario. Because the categories provided by the survey in some cases (e.g., race/ethnicity categories) did not match the categories from the Census, we aggregated some of them to obtain an approximation to the categories in the Census. The aggregation is described in detail in the Appendix.

## Geographical location {#sec-geographical-location}

For each survey participant certain data was automatically captured. This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset. Because our interest lies in exploring vaccination status within a governamental administrative unit that encompasses multiple cities, we grouped the municipalities from the dataset according to the regional government aggregations provided by the Association of Municipalities of Ontario, which groups municipalities within regions, counties, districts, and single-tiers [@ontario-municipalities]. The geographical regions were obtained from the websie of the province of Ontario <https://www.ontario.ca/page/list-ontario-municipalities#section-3>

@@

We used a multivariable logistic regression with the socio-economic factors described in @sec-socio-demographic-factors and the aggregated demographical locations from @sec-geographical-location.

>>>>

---
title: "Covid-19 Fields Survey Data Protocol"
author:
  - name: Ariel Mundo Ortiz
    orcid: 0000-0002-6014-4538
    email: ariel.mundo.ortiz@umontreal.ca
    affiliations:
      - name: École de santé publique - Université de Montréal
        address: 7101 Park Avenue
        city: Montreal
        state: QC
        postal-code: H3N 1X9
date: "`r Sys.Date()`"
format:
  pdf:
    link-citations: true
    number-sections: true
editor: visual
bibliography: bibliography/refs.bib
keep-tex: true
csl: bibliography/jama.csl
knitr:
  opts_chunk:
    dev: "ragg_png"
---

@@

Studies that analyze geographical variations in vaccine uptake can help inform public health decision-makers to design policies to that are aimed at addressing vaccination disparities. In this regard, previous geographical (spatial) analyses of vaccination rates have shown that variations in vaccine uptake can occur within small governmental administrative units (e.g., counties in the case of the US) [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that geographical analyses can be predictive of booster uptake patterns [@wood2022].

In Canada, studies that have used a spatial approach to analyze vaccine uptake have shown disparities in vaccination rates across low and high income neighborhoods in the city of Toronto [@choi2021], among adolescents from deprived neighborhoods in the city of Montreal [@mckinnon2021], and highlighted disparities in vaccination status depending on age, income, and ethnic origin in all of the Canadian provinces[@guay2022]. However, to the best of our knowledge, there are no studies that have analyzed vaccination status within a province to identify inequalities that may exist within these geographical areas. A dissagregated view of variability within a province can help understand the barriers for vaccine delivery in the case of visible minorities, which have been disproportionately impacted in the pandemic [@hussain2022].

@@

This study will examine self-reported COVID-19 vaccination status in the province of Ontario in order to determine the influence in vaccination status due to socio-economic (e.g., ethnic origin, age, income) and geographical factors (at level of the Health Regions of Ontario).

@@

| Variable             | Values                                                                                                                          |
|:----------------|-------------------------------------------------------|
| Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
| Income bracket (CAD) | \<15,000, 15,000-24,999, 25,000-39,999, 40,000-59,999, 60,000-89,999, \>90,000                                                  |
| Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, other |
                                                                                                                      |

@@

From the survey, we selected the question regarding vaccination status:

@@

<!-- -   (If answered "yes" to the previous question) "Have you received the second dose of the COVID vaccine?" with possible answers "yes" and "no" -->

@@

The original dataset contained 39,029 entries (where each entry corresponded to a set of answers provided by a unique respondent). Following a preliminary analysis to identify the missing rates across the different answers within each entry, it was identified that many of the answers had high missing rates (\>80%) (note that the graph will be in the Appendix). Therefore, the dataset was cleaned in order to contain only the independent variables of interest with the lowest missing rates (@tbl-covariates) and the dependent variable.

@@

### Geographical location {#sec-geographical-location}

For each survey participant certain data was automatically captured. This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset. Because our interest was to analyze the differences between Health Regions, we assigned the location (city) of each entry to its correspondent Health Region following a multi-step process: First, we used the municipality information from the website of the Association of Municipalities of Ontario (AMO) to assign geographical regions to each entry on the dataset. Secondly, we used a publicly available dataset of long-term care homes by Local Health Integrated Network (LHIN), which were the geographical divisions for health in Ontario before the adoption of the Health Regions, to match each city with its corresponding LHIN. Cities that did not have a LHIN entry on the dataset were searched in the websites of the LHINs and manually added to the dataset. Finally, we assigned the corresponding Health Regio to each city by matching the LHINs to the Health Regions that now encompass them, following the information provided by Ontario Health (details of the complete process are in the Appendix).

@@

We identified differences between the proportions of various socio-economic factors the survey respondents when compared to the 2016 Census data for Ontario. These factors included age groups, income, and ethnicity/race identified. Additionally, because the Census divisions do not match the exact boundaries of the Census, we also obtained population estimates for each Health Region from the Ontario Health website in order to correct for the population within each Health Region. We corrected for all of these factors (age group, race/ethnicity, income, Health Region Population) using an iterative proportional fitting procedure (also known as *raking*) [@deming1940] in R using the `survey` package. The proportions of each of these factors and the data from the 2016 Census Data for Ontario used for the corrections can be found in the Appendix. Of notice, because the categories provided by the survey in some cases (e.g., race/ethnicity categories) did not match the categories from the Census, we aggregated them where appropriate to obtain an approximation to the categories in the Census.

<!-- ## Geographical location {#sec-geographical-location} -->

<!-- For each survey participant certain data was automatically captured. This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset. Due to our interest in exploring vaccination status within a health region that encompasses multiple cities, we grouped the municipalities from the dataset according to the regional government aggregations provided by the Association of Municipalities of Ontario, which groups municipalities within regions, counties, districts, and single-tiers [@ontario-municipalities]. The geographical regions were obtained from the websie of the province of Ontario <https://www.ontario.ca/page/list-ontario-municipalities#section-3> -->

@@

We used a fixed-effects logistic regression to estimate the probability of vaccination depending on the socio-economic factors described in @sec-socio-demographic-factors and the Health Regions from @sec-geographical-location. We examined the interaction of income and Health Region as it has been shown that there has been an increase in income inequality within the provinces of Canada over time [@marchand2020], and the relatively recent implementation of the Health Regions motivated us to explore if income disparity may be an important factor to consider within each of the Health Regions. Our model appears in @eq-model.

$$
\log \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}}= \beta_0+ \beta_{1}\textrm{(age group)} +\beta_{2} \textrm{ race} + \beta_3 \textrm{(Health Region} \times \textrm{income)} 
$$ {#eq-model}

Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine.

# Results

## Descriptive statistics of the Fields Covid-19 Survey

@tbl-descriptive-stats shows the descriptive statistics (uncorrected) from the Fields Covid-19 survey data for vaccination status and each of the covariates analyzed.

```{r, load-prepare-data}
#| echo: false
#| warning: false
#| message: false

#load packages

library(weights)
library(here)
library(survey)
library(magrittr)
library(tidyverse)
library(lme4)
library(gtsummary)

data<-read.csv(here("data","clean_dataset.csv"))

# remove rows with missing observations for race

clean_data<-data%>%drop_na(race)

#remove "municipal status" as is not part of the analysis (to avoid unexpected errors)

clean_data<-clean_data%>%
  select(-Municipal.status)

#make variables of interest factors
clean_data<-clean_data %<>%
  mutate_at(c("income_ord","race","age_group","location","Geographic_area","Geographic_area_title"),factor)
# recode the response of interest for categorical analysis
clean_data<-clean_data %>%
  mutate(first_dose_m=case_when(
    first_dose=="yes"~1,
    first_dose=="no"~0
  ))

# get percentages to see how different the data is from the Census (details of the Census values are in
# Data_cleaning_and_covariates_Jan_05.qmd in directory "data cleaning")

##### Summary of the dataset####

data_summary<-clean_data %>%
  tbl_summary(percent="row",
    by= first_dose,
    include=c(income_ord,
    age_group,
    Health_Region,
    race)
  ) %>% bold_labels()

```

```{r,summary-table}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics of the Fields Covid-19 Survey"
#| message: false
#| echo: false

data_summary

```

### 2023-01-13 protocol/Background_and_methods_Jan_12.tex
updated protocol

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Covid Fields Data Protocol},
  pdfauthor={Ariel Mundo Ortiz},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Covid Fields Data Protocol}
\author{Ariel Mundo Ortiz}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, breakable, sharp corners, frame hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

Studies that analyze geographical variations in vaccine uptake can help
inform public health decision-makers to design policies to that are
aimed at addressing vaccination disparities. In this regad, previous
geographical (spatial) analyses of vaccination rates have shown that
variations in vaccine uptake can occur within small governamental
administrative units (e.g., counties in the case of the
US)\textsuperscript{\protect\hyperlink{ref-mollalo2021}{12}--\protect\hyperlink{ref-bhuiyan2022}{15}},
and that geographical analyses can be predictive of booster uptake
patterns\textsuperscript{\protect\hyperlink{ref-wood2022}{16}}.

In Canada, studies that have used a spatial approach to analyze vaccine
uptake have shown disparities in vaccination rates across low and high
income neighborhoods in the city of
Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{17}}, among
adolescents from deprived neighborhoods in the city of
Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{18}}, and
highlighted disparities in vaccination status depending on age, income,
and ethnic origin in all of the Canadian
provinces\textsuperscript{\protect\hyperlink{ref-guay2022}{8}}. However,
to the best of our knowledge, there are no studies that have analyzed
vaccination status within a province at the district/rural municipality
level, which can be useful to identify inequalities over these
geographical areas, thus providing a dissagregated view that can help
understand the barriers for vaccine delivery in the case of visible
minorities, which have been disproportionately impacted in the
pandemic\textsuperscript{\protect\hyperlink{ref-hussain2022}{19}}.

@@

\hypertarget{tbl-covariates}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2083}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7917}}@{}}
\caption{\label{tbl-covariates}Selected socio-economic factors from the
survey}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Values
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Values
\end{minipage} \\
\midrule()
\endhead
Age group & 15-24,25-34,35-44,45-54,55-64, 65+ \\
Income bracket (CAD) & \textless15,000, 15,000-24,999, 25,000-39,999,
40,000-59,999, 60,000-89,999, \textgreater90,000 \\
Race/ethnicity & Arab/Middle Eastern, Black, East Asian/Pacific
Islander, Indigenous, Latin American, Mixed, South Asian, White
Caucasian, other \\
Employment status & yes, no \\
\bottomrule()
\end{longtable}

@@

We selected two of the questions that were asked in the survey regarding
vaccination status:

\begin{itemize}
\item
  ``Have you received the first dose of the COVID vaccine?'', with
  possible answers ``yes'' and ``no''
\item
  (If answered ``yes'' to the previous question) ``Have you received the
  second dose of the COVID vaccine?'' with possible answers ``yes'' and
  ``no''
\end{itemize}

@@

The original dataset contained 39,029 entries (where each entry
corresponded to a set of answers provided by a unique respondent).
Following a preliminary analysis to identify the missing rates across
the different answers within each entry, it was identified that many of
the answers had high missing rates (\textgreater80\%) (note that the
graph will be in the Appendix). Therefore, the dataset was cleaned in
order to contain only the independent variables of interest with the
lowest missing rates (Table~\ref{tbl-covariates}) and the dependent
variables.

@@

@@

Differences between the clean dataset and the 2016 Census data for
Ontario regarding proportions of age groups, income, and ethnicity/race
were identified. Therefore, the proportions of each of these variables
were corrected using an iterative proportional fitting procedure (also
known as
\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{21}}
in R using the \texttt{survey} package. Proportions for the correction
were obtained from the 2016 Census Data for Ontario. Because the
categories provided by the survey in some cases (e.g., race/ethnicity
categories) did not match the categories from the Census, we aggregated
some of them to obtain an approximation to the categories in the Census.
The aggregation is described in detail in the Appendix.

\hypertarget{sec-geographical-location}{%
\subsection{Geographical location}\label{sec-geographical-location}}

For each survey participant certain data was automatically captured.
This included the nearest municipality, which resulted in a total of 578
different municipalities within the dataset. Because our interest lies
in exploring vaccination status within a governamental administrative
unit that encompasses multiple cities, we grouped the municipalities
from the dataset according to the regional government aggregations
provided by the Association of Municipalities of Ontario, which groups
municipalities within regions, counties, districts, and
single-tiers\textsuperscript{\protect\hyperlink{ref-ontario-municipalities}{22}}.
The geographical regions were obtained from the websie of the province
of Ontario
\url{https://www.ontario.ca/page/list-ontario-municipalities\#section-3}

@@

We used a multivariable logistic regression with the socio-economic
factors described in Section~\ref{sec-socio-demographic-factors} and the
aggregated demographical locations from
Section~\ref{sec-geographical-location}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-ontario-municipalities}{}}%
\CSLLeftMargin{22. }%
\CSLRightInline{Ontario {M}unicipalities. Accessed December 30, 2022.
\url{https://www.amo.on.ca/about-us/municipal-101/ontario-municipalities}}

>>>>

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Covid-19 Fields Survey Data Protocol},
  pdfauthor={Ariel Mundo Ortiz},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Covid-19 Fields Survey Data Protocol}
\author{Ariel Mundo Ortiz}
\date{1/12/23}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, enhanced, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

Studies that analyze geographical variations in vaccine uptake can help
inform public health decision-makers to design policies to that are
aimed at addressing vaccination disparities. In this regard, previous
geographical (spatial) analyses of vaccination rates have shown that
variations in vaccine uptake can occur within small governmental
administrative units (e.g., counties in the case of the
US)\textsuperscript{\protect\hyperlink{ref-mollalo2021}{12}--\protect\hyperlink{ref-bhuiyan2022}{15}},
and that geographical analyses can be predictive of booster uptake
patterns\textsuperscript{\protect\hyperlink{ref-wood2022}{16}}.

In Canada, studies that have used a spatial approach to analyze vaccine
uptake have shown disparities in vaccination rates across low and high
income neighborhoods in the city of
Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{17}}, among
adolescents from deprived neighborhoods in the city of
Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{18}}, and
highlighted disparities in vaccination status depending on age, income,
and ethnic origin in all of the Canadian
provinces\textsuperscript{\protect\hyperlink{ref-guay2022}{8}}. However,
to the best of our knowledge, there are no studies that have analyzed
vaccination status within a province to identify inequalities that may
exist within these geographical areas. A dissagregated view of
variability within a province can help understand the barriers for
vaccine delivery in the case of visible minorities, which have been
disproportionately impacted in the
pandemic\textsuperscript{\protect\hyperlink{ref-hussain2022}{19}}.

@@

\hypertarget{tbl-covariates}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2361}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7639}}@{}}
\caption{\label{tbl-covariates}Selected socio-economic factors from the
survey}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Values
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Values
\end{minipage} \\
\midrule()
\endhead
Age group & 15-24,25-34,35-44,45-54,55-64, 65+ \\
Income bracket (CAD) & \textless15,000, 15,000-24,999, 25,000-39,999,
40,000-59,999, 60,000-89,999, \textgreater90,000 \\
Race/ethnicity & Arab/Middle Eastern, Black, East Asian/Pacific
Islander, Indigenous, Latin American, Mixed, South Asian, White
Caucasian, other \\
& \\
\bottomrule()
\end{longtable}

@@

From the survey, we selected the question regarding vaccination status:

\begin{itemize}
\tightlist
\item
  ``Have you received the first dose of the COVID vaccine?'', with
  possible answers ``yes'' and ``no''
\end{itemize}

@@

The original dataset contained 39,029 entries (where each entry
corresponded to a set of answers provided by a unique respondent).
Following a preliminary analysis to identify the missing rates across
the different answers within each entry, it was identified that many of
the answers had high missing rates (\textgreater80\%) (note that the
graph will be in the Appendix). Therefore, the dataset was cleaned in
order to contain only the independent variables of interest with the
lowest missing rates (Table~\ref{tbl-covariates}) and the dependent
variable.

@@

\hypertarget{sec-geographical-location}{%
\subsubsection{Geographical location}\label{sec-geographical-location}}

For each survey participant certain data was automatically captured.
This included the nearest municipality, which resulted in a total of 578
different municipalities within the dataset. Because our interest was to
analyze the differences between Health Regions, we assigned the location
(city) of each entry to its correspondent Health Region following a
multi-step process: First, we used the municipality information from the
website of the Association of Municipalities of Ontario (AMO) to assign
geographical regions to each entry on the dataset. Secondly, we used a
publicly available dataset of long-term care homes by Local Health
Integrated Network (LHIN), which were the geographical divisions for
health in Ontario before the adoption of the Health Regions, to match
each city with its corresponding LHIN. Cities that did not have a LHIN
entry on the dataset were searched in the websites of the LHINs and
manually added to the dataset. Finally, we assigned the corresponding
Health Regio to each city by matching the LHINs to the Health Regions
that now encompass them, following the information provided by Ontario
Health (details of the complete process are in the Appendix).

@@

We identified differences between the proportions of various
socio-economic factors the survey respondents when compared to the 2016
Census data for Ontario. These factors included age groups, income, and
ethnicity/race identified. Additionally, because the Census divisions do
not match the exact boundaries of the Census, we also obtained
population estimates for each Health Region from the Ontario Health
website in order to correct for the population within each Health
Region. We corrected for all of these factors (age group,
race/ethnicity, income, Health Region Population) using an iterative
proportional fitting procedure (also known as
\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{21}}
in R using the \texttt{survey} package. The proportions of each of these
factors and the data from the 2016 Census Data for Ontario used for the
corrections can be found in the Appendix. Of notice, because the
categories provided by the survey in some cases (e.g., race/ethnicity
categories) did not match the categories from the Census, we aggregated
them where appropriate to obtain an approximation to the categories in
the Census.

@@

We used a fixed-effects logistic regression to estimate the probability
of vaccination depending on the socio-economic factors described in
Section~\ref{sec-socio-demographic-factors} and the Health Regions from
Section~\ref{sec-geographical-location}. We examined the interaction of
income and Health Region as it has been shown that there has been an
increase in income inequality within the provinces of Canada over
time\textsuperscript{\protect\hyperlink{ref-marchand2020}{22}}, and the
relatively recent implementation of the Health Regions motivated us to
explore if income disparity may be an important factor to consider
within each of the Health Regions. Our model appears in
Equation~\ref{eq-model}.

\begin{equation}\protect\hypertarget{eq-model}{}{
\log \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}}= \beta_0+ \beta_{1}\textrm{(age group)} +\beta_{2} \textrm{ race} + \beta_3 \textrm{(Health Region} \times \textrm{income)} 
}\label{eq-model}\end{equation}

Where \(p\textrm{(vac)}\) indicates the probability of having received
the first dose of a Covid-19 vaccine.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{descriptive-statistics-of-the-fields-covid-19-survey}{%
\subsection{Descriptive statistics of the Fields Covid-19
Survey}\label{descriptive-statistics-of-the-fields-covid-19-survey}}

Table~\ref{tbl-descriptive-stats} shows the descriptive statistics
(uncorrected) from the Fields Covid-19 survey data for vaccination
status and each of the covariates analyzed.

\hypertarget{tbl-descriptive-stats}{}
\begin{longtable}[]{@{}lcc@{}}
\caption{\label{tbl-descriptive-stats}Descriptive Statistics of the
Fields Covid-19 Survey}\tabularnewline
\toprule()
\textbf{Characteristic} & \textbf{no}, N = 1,000 & \textbf{yes}, N =
2,709 \\
\midrule()
\endfirsthead
\toprule()
\textbf{Characteristic} & \textbf{no}, N = 1,000 & \textbf{yes}, N =
2,709 \\
\midrule()
\endhead
\textbf{income\_ord} & & \\
15000\_24999 & 167 (34\%) & 326 (66\%) \\
25000\_39999 & 144 (31\%) & 316 (69\%) \\
40000\_59999 & 116 (25\%) & 348 (75\%) \\
60000\_89999 & 71 (17\%) & 343 (83\%) \\
over\_90000 & 251 (25\%) & 752 (75\%) \\
under\_15000 & 251 (29\%) & 624 (71\%) \\
\textbf{age\_group} & & \\
16\_24 & 230 (27\%) & 634 (73\%) \\
25\_34 & 193 (27\%) & 518 (73\%) \\
35\_44 & 136 (25\%) & 400 (75\%) \\
45\_54 & 130 (28\%) & 335 (72\%) \\
55\_64 & 80 (20\%) & 317 (80\%) \\
65\_and\_over & 231 (31\%) & 505 (69\%) \\
\textbf{Health\_Region} & & \\
Central & 224 (28\%) & 581 (72\%) \\
East & 135 (23\%) & 448 (77\%) \\
North East & 36 (34\%) & 71 (66\%) \\
North West & 6 (12\%) & 45 (88\%) \\
Toronto & 371 (28\%) & 953 (72\%) \\
West & 228 (27\%) & 611 (73\%) \\
\textbf{race} & & \\
arab\_middle\_eastern & 77 (35\%) & 141 (65\%) \\
black & 120 (39\%) & 186 (61\%) \\
east\_asian\_pacific\_islander & 72 (23\%) & 236 (77\%) \\
indigenous & 82 (38\%) & 134 (62\%) \\
latin\_american & 70 (38\%) & 112 (62\%) \\
mixed & 111 (34\%) & 213 (66\%) \\
other & 134 (34\%) & 255 (66\%) \\
south\_asian & 89 (23\%) & 294 (77\%) \\
white\_caucasian & 245 (18\%) & 1,138 (82\%) \\
\bottomrule()
\end{longtable}

@@

\leavevmode\vadjust pre{\hypertarget{ref-marchand2020}{}}%
\CSLLeftMargin{22. }%
\CSLRightInline{Marchand Y, Dubé J, Breau S. Exploring the causes and
consequences of regional income inequality in canada. \emph{Economic
Geography}. 2020;96(2):83-107.
doi:\href{https://doi.org/10.1080/00130095.2020.1715793}{10.1080/00130095.2020.1715793}}
